---
title: "**The rOpenSpain project**"
subtitle: "R and open data"
author: "**Iñaki Úcar** | Postdoctoral Fellow @ uc3m-Santander Big Data Institute"
abstract: |
  | rOpenSpain is a community of R, open data and reproducibility enthusiasts
  | that gathers and organizes to create R packages of the highest quality for
  | the exploitation of Spanish data of general interest. In this talk we will
  | review its (brief) history, organization and the subprojects that comprise it.
date: "June 9, 2022"
output:
  xaringan::moon_reader:
    self_contained: false
    css: [default, rOpenSpain.css, extra.css]
    nature:
      beforeInit: "rOpenSpain.js"
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  out.width = "100%",
  cache = TRUE,
  comment = "#>",
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  hiline = TRUE,
  dpi = 300
)

pkgs.maps <- c("mapSpain", "LAU2boundaries4spain", "CatastRo", "CatastRoNav", "caRtociudad")
pkgs.government <- c("BOE", "infoelectoral", "senadoRES", "opendataes")
pkgs.statistics <- c("istacbaser", "MicroDatosEs", "MorbiditySpainR", "Siane")
pkgs.climate <- c("climaemet", "airqualityES")
pkgs.economy <- c("tidyBdE")
pkgs.utils <- c("spanish")
```

class: base28

# Introduction

.left-column[
## When
]

.right-column[
It was **February 2018** when four people came together around open data:

- **Carlos J. Gil Bellosta**, statistical consultant, CEO @ circiteR
- **Luz Frías**, data developer, CTO @ circiteR
- **José Manuel Vera**, senior data scientist
- **Iñaki Úcar**, postdoctoral fellow @ IBiDat

### Motto

> rOpenSci is our form; Spanish public data, our matter

]

---
class: base28

# Introduction

.left-column[
## When
## What
]

.right-column[
The **github.com/rOpenSpain** organization
- Website (ropenspain.es)
- Onboarding info
- Templates (shout-outs to Diego Hernangómez)
- **Packages**

The **ropenspain.slack.com** channel
- Do not hesitate to contact us for an invitation!
]

---
class: base28

# Introduction

.left-column[
## When
## What
## How
]

.right-column[
Do you have a package about Spanish data? <br>**Bring it to rOpenSpain!**
- The author transfers the repo to our GH organization
- The author retains full admin rights
- The package is added to the webpage and is available for installation
  through our [r-universe](https://ropenspain.r-universe.dev) organization
- We try to encourage R packaging best practices
- We try to help each other out

Do you need help? Get an invite to our Slack!
]

---

# Introduction

.left-column[
## When
## What
## How
## Overview
]

.right-column[
.pull-left[
### Statistical data
- istacbaser
- MicroDatosES
- MorbiditySpainR
- Siane

### Maps
- mapSpain
- LAU2boundaries4spain
- CatastRo, CatastRoNav
- caRtociudad
]

.pull-right[
### Government
- BOE
- infoelectoral
- senadoRES
- opendataes

### Economy
- tidyBdE

### Climate
- climaemet
- airqualityES
]
]

---
class: inverse, center, middle

# Statistical data

```{r statistics, child=paste0("packages/", pkgs.statistics, ".Rmd")}
```

---
class: inverse, center, middle

# Maps

```{r maps, child=paste0("packages/", pkgs.maps, ".Rmd")}
```

---
class: inverse, center, middle

# Government

```{r government, child=paste0("packages/", pkgs.government, ".Rmd")}
```

---
class: inverse, center, middle

# Economy

```{r economy, child=paste0("packages/", pkgs.economy, ".Rmd")}
```

---
class: inverse, center, middle

# Climate

```{r climate, child=paste0("packages/", pkgs.climate, ".Rmd")}
```

---
class: base24

# Join rOpenSpain!

- Much more could be done about the retrieval of statistical data
  - In particular, we still lack an **"inebaser"** similar to istacbaser

- Mapping needs are pretty much covered, but improvements can always be made
  - We still lack cadastral access to the Basque Country
  - Others would require a transition to the new spatial stack (based on **sf**)

- **opendataes** needs your help!
  - It does a great job setting the framework required to work with datos.gob.es
  - There are many publishers and formats that could be added

- Do you have other ideas? We'd love to hear you out!

---
class: center, middle, end-slide

# Thanks!

Slides created via the R package [**xaringan**](https://github.com/yihui/xaringan) with the [**rosxaringan**](https://github.com/rOpenSpain/rosxaringan) template.

The chakra comes from [remark.js](https://remarkjs.com), [**knitr**](https://yihui.org/knitr/), and [R Markdown](https://rmarkdown.rstudio.com).
